---
title: "Get online and free to use Copernicus data"
author: "Agathe Dumont"
output: html_document
---
Data source:
https://cds.climate.copernicus.eu/cdsapp#!/dataset/sis-agrometeorological-indicators?tab=doc

resolution : 0.1° x 0.1° (10 km environs), from 1979 to present

Citation:
Boogaard, H., Schubert, J., De Wit, A., Lazebnik, J., Hutjes, R., Van der Grijn, G., (2020): Agrometeorological indicators from 1979 to present derived from reanalysis. Copernicus Climate Change Service (C3S) Climate Data Store (CDS). (Accessed on DD-MMM-YYYY), 10.24381/cds.6c68c9bb
DOI:10.24381/cds.6c68c9bb

# Data
```{r}
# data C4F
C4F_data = readRDS(paste0("../",data_path))
# api copenricus data store
wf_set_key(user = user_id, key = cds_key)

# day, month, year
una12= c("01","02","03","04","05","06","07","08","09","10","11","12")
una31 = c(una12,as.character(seq(from = 13, to = 31, by = 1)))
fichier_concat = "dataclim_2020212223_CI.nc"
year = c("2018","2019","2020","2021","2022","2023")
year_date = as.Date(c("01/01/18","01/01/19","01/01/20","01/01/21","01/01/22","01/01/23"), "%m/%d/%y")

# CI box 
zone = c(floor(max(C4F_data$latitude)) + 1, # zone max latitude
         floor(min(C4F_data$longitude))-1, # zone min longitude
         floor(min(C4F_data$latitude))-1, # zone min latitude
         floor(max(C4F_data$longitude))+1 ) # zone max longitude

```

# Extract climat data from copernicus to your folder
```{r eval=FALSE, include=FALSE}

list_va_clim = c("2m_temperature","10m_wind_speed","precipitation_flux")
va_humidity = "2m_relative_humidity"

# For each year
for(y in year) { # for year in the vector year
  for(v in list_va_clim) # for each clim variable
  extract_clim_va(
    Va = v,
    Year = y,
    Month = una12,
    download_path = path_data_raw_zip,
    Stat = "24_hour_mean",
    Day = una31,
    Zone_chr = "CI",
    Zone = zone,
    user_id = user_id
  )
}

  
  extract_clim_va(
    Va = "2m_relative_humidity",
    Year = y,
    Month = una12,
    download_path = path_data_raw_zip,
    Stat = "24_hour_mean",
    Day = una31,
    Zone_chr = "CI",
    Zone = zone,
    user_id = user_id,
     Time = c("15_00")
  )
}

```

Attention : unzip the file by hand ...

# Create a ncdf file
```{r}

# create ncdf file empty from an other
dossier_nc  = "agrometeorological_copernicus_2018_2m_temperature_CI"
dossier_path <- paste("../output",dossier_nc,sep = "")
fichier_list = list.files(dossier_path)
nc_file <- nc_open(paste(dossier_path,"/",fichier_list[1],sep = ""))
temp_dim =  ncvar_get(nc_file,"time")

## dimension 
dim_state = list(
lon_dim = 
  ncdim_def(
    nc_file$var[[1]]$dim[[1]]$name,
    units = nc_file$var[[1]]$dim[[1]]$units,
    nc_file$var[[1]]$dim[[1]]$vals
  ),

lat_dim = 
  ncdim_def(
    nc_file$var[[1]]$dim[[2]]$name,
    units = nc_file$var[[1]]$dim[[2]]$units,
    nc_file$var[[1]]$dim[[2]]$vals
  ),

time_dim = 
  ncdim_def(
    nc_file$var[[1]]$dim[[3]]$name,
    units = nc_file$var[[1]]$dim[[3]]$units,
    seq(from = temp_dim, to = temp_dim + sum(yearDays(year_date)) -1)
  )
)

nc_close(nc_file)

## get variables of ncdf files
hum_15h_va = nc_var_def(path_data_raw, dossier_nc = "agrometeorological_copernicus_2018_2m_relative_humidity_15_00_CI")
temp_va = nc_var_def(path_data_raw, dossier_nc = "agrometeorological_copernicus_2018_2m_temperature_CI")
cloud_va = nc_var_def(path_data_raw, dossier_nc = "agrometeorological_copernicus_2018_10m_wind_speed_CI")
wind_va = nc_var_def(path_data_raw, dossier_nc = "agrometeorological_copernicus_2018_cloud_cover_CI")
prec_va = nc_var_def(path_data_raw, dossier_nc = "agrometeorological_copernicus_2018_precipitation_flux_CI")


## create an ncdf file with all clim variables and all days
new_ncdf = nc_create(
  "../output/meteo_copernicus_download.nc",
  vars = list(hum_9h_va,hum_15h_va,temp_va,cloud_va,wind_va,prec_va)
)

nc_close(new_ncdf)

```

# Concatenation
```{r}
# list of file to concat
list_to_concat = grep("^agro",list.files("../output/"),value = T)

# concat all the files on one ncdf file
for (j in list_to_concat) {
  va_concat_nc_function(
    fichier_concat =  "../output/meteo_copernicus_download.nc",
    Path_data_raw = "../output/",
    fichiers_to_concat = j
  )
}
```

# Position of each observation in one grid cell
```{r}
# grid resolution
resolution = 0.10

# Take latitude and longitude 
data_clim_concat = nc_open("../output/meteo_copernicus_download.nc")
long = ncvar_get(data_clim_concat, "lon")
lat = ncvar_get(data_clim_concat, "lat")
nc_close(data_clim_concat)
range_lon = range(long)
range_lat = range(lat)

#  Each plot is assigned to a grid cell. 
cut_longitude = cut(
  C4F_data$longitude,
  breaks = seq(range_lon[1] - resolution / 2,
               range_lon[2] + resolution / 2,
               resolution)
)

cut_latitude = cut(
  C4F_data$latitude,
  breaks = seq(range_lat[1] - resolution / 2,
               range_lat[2] + resolution /2,
               resolution)
)

# center of the cell
cut_longitude = as.character(cut_longitude)
cut_longitude = unlist(strsplit(cut_longitude, ","))[c(T, F)]
cut_longitude = as.numeric(sub("\\(", "", cut_longitude)) + resolution / 2
cut_latitude = as.character(cut_latitude)
cut_latitude = unlist(strsplit(cut_latitude, ","))[c(T, F)]
cut_latitude = as.numeric(sub("\\(", "", cut_latitude)) + resolution / 2

# Add variable to dataframe
C4F_data$cut_lon = cut_longitude
C4F_data$cut_lat = cut_latitude
```

# Get variable by year
```{r}
list_year = c(2018:2022)
dataclimnc = nc_open("../output/meteo_copernicus_download.nc")
nc_close(dataclimnc)
list_va_clim = names(dataclimnc$var)

by_years = lapply(list_year, function(y) {
  byyear = lapply(list_va_clim, function(x) {
    add_vaclim_to_dataframe_byyear(
      va = x,
      data = C4F_data,
      path_data_clim = "../output/meteo_copernicus_download.nc",
      absolute_time = T,
      year = y
    )
  })
  df_byyear = base::do.call(cbind, byyear)
})

df_byyears = base::do.call(cbind, by_years)
```

# Get variable by saison
```{r}
## dry large
drylarge_years = lapply(list_year, function(y) {
  drylarge_year = lapply(list_va_clim, function(x) {
    add_vaclim_to_dataframe_byseason(
      data = C4F_data,
      path_data_clim = "../output/meteo_copernicus_download.nc",
      va = x,
      month_s = c(12, 1, 2, 3, 4),
      season_name = "drylarge",
      absolute_time = T,
      year = y
    )
    
  })
  df_drylargeyears = base::do.call(cbind, drylarge_year)
})

df_drylarge_years =  base::do.call(cbind, drylarge_years)

## dry small
drysmall_years = lapply(list_year, function(y) {
  drysmall_year = lapply(list_va_clim, function(x) {
    add_vaclim_to_dataframe_byseason(
      data = C4F_data,
      path_data_clim = "../output/meteo_copernicus_download.nc",
      va = x,
      month_s = c(8,9),
      season_name = "drysmall",
      absolute_time = T,
      year = y
    )
    
  })
  df_drysmallyears = base::do.call(cbind, drysmall_year)
})

df_drysmall_years =  base::do.call(cbind, drysmall_years)

## rain large
rainlarge_years = lapply(list_year, function(y) {
  rainlarge_year = lapply(list_va_clim, function(x) {
    add_vaclim_to_dataframe_byseason(
      data = C4F_data,
      path_data_clim = "../output/meteo_copernicus_download.nc",
      va = x,
      month_s = c(5,6,7),
      season_name = "rainlarge",
      absolute_time = T,
      year = y
    )
    
  })
  df_rainlargeyears = base::do.call(cbind, rainlarge_year)
})

df_rainlarge_years =  base::do.call(cbind, rainlarge_years)

## rain small
rainsmall_years = lapply(list_year, function(y) {
  rainsmall_year = lapply(list_va_clim, function(x) {
    add_vaclim_to_dataframe_byseason(
      data = C4F_data              ,
      path_data_clim = "../output/meteo_copernicus_download.nc",
      va = x,
      month_s = c(10,11),
      season_name = "rainsmall",
      absolute_time = T,
      year = y
    )
    
  })
  df_rainsmallyears = base::do.call(cbind, rainsmall_year)
})

df_rainsmall_years =  base::do.call(cbind, rainsmall_years)
```

# Create data set
```{r}
data_meteo_absolute = cbind(
  df_byyears,
  df_drylarge_years,
  df_drysmall_years,
  df_rainlarge_years,
  df_rainsmall_years
)

data_meteo_absolute = as.data.frame(data_meteo_absolute)

data_meteo_absolute$latitude = C4F_data$latitude
data_meteo_absolute$longitude = C4F_data$longitude
data_meteo_absolute$date = C4F_data$date
data_meteo_absolute$code_parcelle = C4F_data$code_parcelle
```

# Save data set
```{r}
saveRDS(data_meteo_absolute,file = "../output/dataadd_meteoabs.RDS")
```
